What makes listeners infer teaching intentions?

### 2022-01-31 paper/manuscript.Rmd
update methods
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

## Apparatus
The experiment was programmed in Python 3.8.2 using the PsychoPy library (2020.2.4; https://www.psychopy.org/) on a Mac Book Pro with Mac OS X Catalina 10.15.6. Stimuli were played using the Mido library (1.2.9; https://mido.readthedocs.io/en/latest/) on a Max/MSP patcher (8.1.7; https://cycling74.com/products/max). During the experiment, participants listened to the stimuli via headphones (Audio-Technica ATH-M50X).

@@

For the current experiment, 96 performances were chosen from the valid performances in Tominaga et al. (*submitted*). We randomly sampled 24 articulation performances and 24 dynamics performances from the teaching condition as well as 24 articulation performances and 24 dynamics performances from the performing condition. It is important to note that the performances from the teaching condition did not necessarily exhibit specific features of teaching performance that we found in our previous experiments since we randomly sampled the recordings.

@@

Stimuli (performances) were quantified with regard to tempo (interonset intervals; IOIs), articulation (key-overlap time; KOT), dynamics (key velocity; KV) and dynamics contrast (key velocity difference; KV-Diff) only for 16th notes. Interonset intervals are time intervals between onsets of adjacent notes. Larger IOIs indicate slower tempo while smaller IOIs indicate faster tempo. Key-overlap time is the time overlap between two adjacent notes, namely the difference between the offset time of the current note and the onset time of the ensuing note [e.g., @bresin_2000]. Positive KOT values indicate legato styles whereas negative KOT values indicate staccato styles. Key velocity is obtained from MIDI data to describe how fast a performer hit the key. Larger KV values indicate forte styles while smaller KV values indicate piano styles. Additionally, we also measured dynamics contrast where one subcomponent of the technique moves to the other (e.g., from forte to piano) to illustrate how much dynamics contrast a performer made at transition points.

First, we investigated data by looking at correlations between performance features (i.e., IOIs, KOT, KV, KV-Diff) and participants' judgments as "teaching" (i.e., what percentage of participants responded as "yes (teaching)") for each stimulus. Second, we run multiple regression and entered all four performance features to examine which predictor significantly contributed to participants' judgments as teaching. Since articulation and dynamics were comprised of two opposite directional values (i.e., legato vs. staccato, forte vs. piano), we entered each subcomponent (i.e., legato, staccato, forte, piano) separately to represent either an articulation or dynamics feature of the performance for a regression model. For articulation recordings, there were two models. The first model considered only legato parts of the stimuli for KOT and KV, and transition points from legato to staccato for KV-Diff. The second model considered only staccato parts of the stimuli for KOT and KV, and transition points from staccato to legato for KV-Diff. For dynamics recordings, there were two models. The first model considered only forte parts of the stimuli for KV and KOT, and transition points from forte to piano for KV-Diff. The second model considered only piano parts of the stimuli for KV and KOT, and transition points from piano to forte for KV-Diff. With regard to tempo (IOIs), there was only one value for each technique because tempo was consistent across the performance and was not dependent on subcomponents.

>>>>

## Apparatus
The experiment was programmed in Python 3.8.2 using the PsychoPy Python library (2020.2.4; https://www.psychopy.org/) on a Mac Book Pro with Mac OS X Catalina 10.15.6. Stimuli were played using the Mido Python library (1.2.9; https://mido.readthedocs.io/en/latest/) on a Max/MSP patcher (8.1.7; https://cycling74.com/products/max). During the experiment, participants listened to the stimuli via headphones (Audio-Technica ATH-M50X).

@@

For the current experiment, 96 performances were chosen from the valid performances in Tominaga et al. (*submitted*). We randomly sampled 24 articulation performances and 24 dynamics performances from the teaching condition as well as 24 articulation performances and 24 dynamics performances from the performing condition. It is important to note that the performances from the teaching condition did not necessarily exhibit specific features of teaching that we found in our previous experiments (e.g., exaggeration) since we randomly sampled the performances.

@@

Stimuli were quantified with regard to tempo (interonset intervals; IOIs), articulation (key-overlap time; KOT), dynamics (key velocity; KV) and dynamics contrast (key velocity difference; KV-Diff) only for 16th notes. Interonset intervals are time intervals between onsets of adjacent notes. Larger IOIs indicate slower tempo while smaller IOIs indicate faster tempo. Key-overlap time is the time overlap between two adjacent notes, namely the difference between the offset time of the current note and the onset time of the ensuing note [e.g., @bresin_2000]. Positive KOT values indicate legato styles whereas negative KOT values indicate staccato styles. Key velocity is obtained from MIDI data to describe how fast a performer hit the key. Larger KV values indicate forte styles while smaller KV values indicate piano styles. Additionally, we also measured dynamics contrast where one subcomponent of the technique moves to the other (e.g., from forte to piano) to illustrate how much dynamics contrast a performer made at transition points.

First, we investigated data by looking at correlations between performance features (i.e., IOIs, KOT, KV, KV-Diff) and participants' judgments as "teaching" (i.e., what percentage of participants responded as "yes") for each stimulus. Second, we run multiple regression and entered all four performance features to examine which predictor significantly contributed to participants' judgments as teaching. Since articulation and dynamics were comprised of two opposite directional values (i.e., legato vs. staccato, forte vs. piano), we entered each subcomponent (i.e., legato, staccato, forte, piano) separately to represent either an articulation or dynamics feature of the performance for a regression model.

For articulation recordings, there were two models. The first model considered only legato parts of the stimuli for KOT and KV, and transition points from legato to staccato for KV-Diff. The second model considered only staccato parts of the stimuli for KOT and KV, and transition points from staccato to legato for KV-Diff. For dynamics recordings, there were two models. The first model considered only forte parts of the stimuli for KV and KOT, and transition points from forte to piano for KV-Diff. The second model considered only piano parts of the stimuli for KV and KOT, and transition points from piano to forte for KV-Diff. With regard to tempo (IOIs), there was only one value for each technique because tempo was consistent across the performance and was not dependent on subcomponents. Therefore, we entered the same tempo value for legato and staccato or forte and piano parts.

### 2022-01-31 paper/manuscript.tex
update methods
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

Stimuli (performances) were quantified with regard to tempo (interonset intervals; IOIs), articulation (key-overlap time; KOT), dynamics (key velocity; KV) and dynamics contrast (key velocity difference; KV-Diff) only for 16th notes. Interonset intervals are time intervals between onsets of adjacent notes. Larger IOIs indicate slower tempo while smaller IOIs indicate faster tempo. Key-overlap time is the time overlap between two adjacent notes, namely the difference between the offset time of the current note and the onset time of the ensuing note (e.g., Bresin \& Battel, 2000). Positive KOT values indicate legato styles whereas negative KOT values indicate staccato styles. Key velocity is obtained from MIDI data to describe how fast a performer hit the key. Larger KV values indicate forte styles while smaller KV values indicate piano styles. Additionally, we also measured dynamics contrast where one subcomponent of the technique moves to the other (e.g., from forte to piano) to illustrate how much dynamics contrast a performer made at transition points.

First, we investigated data by looking at correlations between performance features (i.e., IOIs, KOT, KV, KV-Diff) and participants' judgments as ``teaching'' (i.e., what percentage of participants responded as ``yes (teaching)'') for each stimulus. Second, we run multiple regression and entered all four performance features to examine which predictor significantly contributed to participants' judgments as teaching. Since articulation and dynamics were comprised of two opposite directional values (i.e., legato vs.~staccato, forte vs.~piano), we entered each subcomponent (i.e., legato, staccato, forte, piano) separately to represent either an articulation or dynamics feature of the performance for a regression model. For articulation recordings, there were two models. The first model considered only legato parts of the stimuli for KOT and KV, and transition points from legato to staccato for KV-Diff. The second model considered only staccato parts of the stimuli for KOT and KV, and transition points from staccato to legato for KV-Diff. For dynamics recordings, there were two models. The first model considered only forte parts of the stimuli for KV and KOT, and transition points from forte to piano for KV-Diff. The second model considered only piano parts of the stimuli for KV and KOT, and transition points from piano to forte for KV-Diff. With regard to tempo (IOIs), there was only one value for each technique because tempo was consistent across the performance and was not dependent on subcomponents.

>>>>

Stimuli were quantified with regard to tempo (interonset intervals; IOIs), articulation (key-overlap time; KOT), dynamics (key velocity; KV) and dynamics contrast (key velocity difference; KV-Diff) only for 16th notes. Interonset intervals are time intervals between onsets of adjacent notes. Larger IOIs indicate slower tempo while smaller IOIs indicate faster tempo. Key-overlap time is the time overlap between two adjacent notes, namely the difference between the offset time of the current note and the onset time of the ensuing note (e.g., Bresin \& Battel, 2000). Positive KOT values indicate legato styles whereas negative KOT values indicate staccato styles. Key velocity is obtained from MIDI data to describe how fast a performer hit the key. Larger KV values indicate forte styles while smaller KV values indicate piano styles. Additionally, we also measured dynamics contrast where one subcomponent of the technique moves to the other (e.g., from forte to piano) to illustrate how much dynamics contrast a performer made at transition points.

First, we investigated data by looking at correlations between performance features (i.e., IOIs, KOT, KV, KV-Diff) and participants' judgments as ``teaching'' (i.e., what percentage of participants responded as ``yes (teaching)'') for each stimulus. Second, we run multiple regression and entered all four performance features to examine which predictor significantly contributed to participants' judgments as teaching. Since articulation and dynamics were comprised of two opposite directional values (i.e., legato vs.~staccato, forte vs.~piano), we entered each subcomponent (i.e., legato, staccato, forte, piano) separately to represent either an articulation or dynamics feature of the performance for a regression model.

For articulation recordings, there were two models. The first model considered only legato parts of the stimuli for KOT and KV, and transition points from legato to staccato for KV-Diff. The second model considered only staccato parts of the stimuli for KOT and KV, and transition points from staccato to legato for KV-Diff. For dynamics recordings, there were two models. The first model considered only forte parts of the stimuli for KV and KOT, and transition points from forte to piano for KV-Diff. The second model considered only piano parts of the stimuli for KV and KOT, and transition points from piano to forte for KV-Diff. With regard to tempo (IOIs), there was only one value for each technique because tempo was consistent across the performance and was not dependent on subcomponents. Therefore, we entered the same tempo value for legato and staccato or forte and piano parts.
